Fast and robust low-rank approximation for high-dimensional seismic data reconstruction

### 2020-12-25 lr5d/paper.tex
2020年12月25日 星期五 03时25分50秒 CST

\history{Date of publication xxxx 00, 0000, date of current version xxxx 00, 0000.}
\doi{10.1109/ACCESS.2020.DOI}

@@

\corresp{Corresponding author: M. Bai (e-mail: baimin2016@126.com).}
\tfootnote{This research is sponsored by the National Natural Science Foundation of China (Nos. \new{41904110,}41704121), \new{the China Postdoctoral Science Foundation (No. 2019M662567), and the Postdoctoral Science Foundation of Hubei Province.}}

\def\thevol{8}
\def\myyear{2020}

@@

\begin{keywords}
Multidimensional seismic data, \old{lowrank}\new{low-rank} approximation, seismic data processing, \new{seismic reconstruction, matrix completion}
\end{keywords}

\titlepgskip=-15pt

@@

\begin{figure}[htb!]
 \centering
 \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-clean-3d}
   \label{fig:syn5d-clean-3d}}
 \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-noisy-3d}
   \label{fig:syn5d-noisy-3d}}
 \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-decimated-3d}
   \label{fig:syn5d-decimated-3d}}        
  \caption{Common offset gathers for the 5D synthetic example with linear events. (a) Clean data. (b) Noisy data. (c) The observed data with 80\% traces randomly removed.}
  \label{fig:syn5d-clean-3d,syn5d-noisy-3d,syn5d-decimated-3d}
\end{figure}

\begin{figure}[htb!]
  \centering
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-svd-3d}
    \label{fig:syn5d-svd-3d}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-n-svd-3d}
    \label{fig:syn5d-n-svd-3d}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-r-svd-3d}
    \label{fig:syn5d-r-svd-3d}} 
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-lmafit-3d}
    \label{fig:syn5d-lmafit-3d}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-n-lmafit-3d}
    \label{fig:syn5d-n-lmafit-3d}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-r-lmafit-3d}
    \label{fig:syn5d-r-lmafit-3d}}       
   \caption{Comparison of reconstruction performance in common offset gathers for the 5D synthetic example with linear events. (a) Reconstructed data using the SVD-based low-rank approximation method. (b) The removed noise from the noisy data corresponding to (a). (c) The reconstruction error corresponding to (a). (d) Reconstructed data using the SVD-free low-rank approximation method. (e) The removed noise from the noisy data corresponding to (d). (f) The reconstruction error corresponding to (d).}
\label{fig:syn5d-svd-3d,syn5d-n-svd-3d,syn5d-r-svd-3d,syn5d-lmafit-3d,syn5d-n-lmafit-3d,syn5d-r-lmafit-3d}
\end{figure}

\begin{figure}[htb!]
  \centering
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-clean-3d-1}
    \label{fig:syn5d-clean-3d-1}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-noisy-3d-1}
    \label{fig:syn5d-noisy-3d-1}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-decimated-3d-1}
    \label{fig:syn5d-decimated-3d-1}}        
   \caption{Common midpoint gathers for the 5D synthetic example with linear events. (a) Clean data. (b) Noisy data. (c) The observed data with 80\% traces randomly removed.}
   \label{fig:syn5d-clean-3d-1,syn5d-noisy-3d-1,syn5d-decimated-3d-1}
\end{figure}

\begin{figure}[htb!]
  \centering
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-svd-3d-1}
    \label{fig:syn5d-svd-3d-1}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-n-svd-3d-1}
    \label{fig:syn5d-n-svd-3d-1}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-r-svd-3d-1}
    \label{fig:syn5d-r-svd-3d-1}} 
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-lmafit-3d-1}
    \label{fig:syn5d-lmafit-3d-1}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-n-lmafit-3d-1}
    \label{fig:syn5d-n-lmafit-3d-1}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-r-lmafit-3d-1}
    \label{fig:syn5d-r-lmafit-3d-1}}       
   \caption{Comparison of reconstruction performance in common midpoint gathers for the 5D synthetic example with linear events. (a) Reconstructed data using the SVD-based low-rank approximation method. (b) The removed noise from the noisy data corresponding to (a). (c) The reconstruction error corresponding to (a). (d) Reconstructed data using the SVD-free low-rank approximation method. (e) The removed noise from the noisy data corresponding to (d). (f) The reconstruction error corresponding to (d).}
\label{fig:syn5d-svd-3d-1,syn5d-n-svd-3d-1,syn5d-r-svd-3d-1,syn5d-lmafit-3d-1,syn5d-n-lmafit-3d-1,syn5d-r-lmafit-3d-1}
\end{figure}

\begin{figure}[htb!]
  \centering
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-clean-2d}
    \label{fig:syn5d-clean-2d}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-noisy-2d}
    \label{fig:syn5d-noisy-2d}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-decimated-2d}
    \label{fig:syn5d-decimated-2d}} 
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-svd-2d}
    \label{fig:syn5d-svd-2d}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{linear/Fig/syn5d-lmafit-2d}
    \label{fig:syn5d-lmafit-2d}}    
   \caption{Comparison of reconstruction performance in a 2D section. (a) Clean data. (b) Noisy data. (c) The observed data. (d) Reconstructed data using the SVD-based low-rank approximation method. (e) Reconstructed data using the SVD-free low-rank approximation method.}
\label{fig:syn5d-clean-2d,syn5d-noisy-2d,syn5d-decimated-2d,syn5d-svd-2d,syn5d-lmafit-2d}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=0.45\textwidth]{linear_time/Fig/synth-ss}
   \caption{Comparison of the middle trace amplitude of each cubes in different datasets. The black line denotes the trace from the clean data.  The red line denotes the noisy data. The observed trace cannot be seen from this FIGURE because it is a blank trace in the selection position. The SVD-based and SVD-free methods are denoted as blue and green lines, respectively.}
   \label{fig:synth-ss}
\end{figure}

\begin{figure}[htb!]
  \centering
  \subfigure[]{\includegraphics[width=0.45\columnwidth]{linear/Fig/simi-syn5d-noisy-2d}
    \label{fig:simi-syn5d-noisy-2d}}
  \subfigure[]{\includegraphics[width=0.45\columnwidth]{linear/Fig/simi-syn5d-decimated-2d}
    \label{fig:simi-syn5d-decimated-2d}} \\
  \subfigure[]{\includegraphics[width=0.45\columnwidth]{linear/Fig/simi-syn5d-svd-2d}
    \label{fig:simi-syn5d-svd-2d}} 
  \subfigure[]{\includegraphics[width=0.45\columnwidth]{linear/Fig/simi-syn5d-lmafit-2d}
    \label{fig:simi-syn5d-lmafit-2d}}  
   \caption{Comparison of reconstruction performance in terms of local similarity. (a) Local similarity of the noisy data. (b) Local similarity of the observed data. (c) Local similarity of the reconstructed data using SVD-based low-rank approximation method. (d) Local similarity of the reconstructed data using SVD-free low-rank approximation method.}
\label{fig:simi-syn5d-noisy-2d,simi-syn5d-decimated-2d,simi-syn5d-svd-2d,simi-syn5d-lmafit-2d}
\end{figure}

\begin{figure}[htb!]
  \centering
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{curve/Fig/curve5d-clean-3d-1}
    \label{fig:curve5d-clean-3d-1}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{curve/Fig/curve5d-noisy-3d-1}
    \label{fig:curve5d-noisy-3d-1}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{curve/Fig/curve5d-decimated-3d-1}
    \label{fig:curve5d-decimated-3d-1}}        
   \caption{Common midpoint gathers for the 5D synthetic example with hyperbolic events. (a) Clean data. (b) Noisy data. (c) The observed data with 80\% traces randomly removed.}
   \label{fig:curve5d-clean-3d-1,curve5d-noisy-3d-1,curve5d-decimated-3d-1}
\end{figure}

\begin{figure}[htb!]
  \centering
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{curve/Fig/curve5d-svd-3d-1}
    \label{fig:curve5d-svd-3d-1}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{curve/Fig/curve5d-n-svd-3d-1}
    \label{fig:curve5d-n-svd-3d-1}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{curve/Fig/curve5d-r-svd-3d-1}
    \label{fig:curve5d-r-svd-3d-1}} 
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{curve/Fig/curve5d-lmafit-3d-1}
    \label{fig:curve5d-lmafit-3d-1}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{curve/Fig/curve5d-n-lmafit-3d-1}
    \label{fig:curve5d-n-lmafit-3d-1}}
  \subfigure[]{\includegraphics[width=0.3\columnwidth]{curve/Fig/curve5d-r-lmafit-3d-1}
    \label{fig:curve5d-r-lmafit-3d-1}}       
   \caption{Comparison of reconstruction performance in common midpoint gathers for the 5D synthetic example with hyperbolic events. (a) Reconstructed data using the SVD-based low-rank approximation method. (b) The removed noise from the noisy data corresponding to (a). (c) The reconstruction error corresponding to (a). (d) Reconstructed data using the SVD-free low-rank approximation method. (e) The removed noise from the noisy data corresponding to (d). (f) The reconstruction error corresponding to (d).}
\label{fig:curve5d-svd-3d-1,curve5d-n-svd-3d-1,curve5d-r-svd-3d-1,curve5d-lmafit-3d-1,curve5d-n-lmafit-3d-1,curve5d-r-lmafit-3d-1}
\end{figure}

\begin{figure}[htb!]
  \centering
  \subfigure[]{\includegraphics[width=0.45\columnwidth]{curve/Fig/simi-curve5d-noisy-3d-1}
    \label{fig:simi-curve5d-noisy-3d-1}}
 \subfigure[]{\includegraphics[width=0.45\columnwidth]{curve/Fig/simi-curve5d-decimated-3d-1}
    \label{fig:simi-curve5d-decimated-3d-1}} \\
  \subfigure[]{\includegraphics[width=0.45\columnwidth]{curve/Fig/simi-curve5d-svd-3d-1}
    \label{fig:simi-curve5d-svd-3d-1}}
  \subfigure[]{\includegraphics[width=0.45\columnwidth]{curve/Fig/simi-curve5d-lmafit-3d-1}
    \label{fig:simi-curve5d-lmafit-3d-1} }
   \caption{Comparison of reconstruction performance in terms of local similarity. (a) Local similarity of the noisy data. (b) Local similarity of the observed data. (c) Local similarity of the reconstructed data using SVD-based low-rank approximation method. (d) Local similarity of the reconstructed data using SVD-free low-rank approximation method.}
\label{fig:simi-curve5d-noisy-3d-1,simi-curve5d-decimated-3d-1,simi-curve5d-svd-3d-1,simi-curve5d-lmafit-3d-1}
\end{figure}

\begin{figure*}[htb!]
  \centering
  \subfigure[]{\includegraphics[width=\textwidth]{field/Fig/field5d-decimated2d-0}
    \label{fig:field5d-decimated2d-0}}
 \subfigure[]{\includegraphics[width=\textwidth]{field/Fig/field5d-svd2d-0}
    \label{fig:field5d-svd2d-0}}
  \subfigure[]{\includegraphics[width=\textwidth]{field/Fig/field5d-lmafit2d-0}
    \label{fig:field5d-lmafit2d-0}}
   \caption{Comparison of reconstruction performance in common offset gathers for the 5D field data example. (a) Observed data with roughly 80\% traces missing. (b) Reconstructed data using the SVD-based low-rank approximation method. (c) Reconstructed data using the SVD-free low-rank approximation method. The frame box areas are zoomed and highlighted in FIGURE \ref{fig:z-decimated,z-svd,z-lmafit}.}
\label{fig:field5d-decimated2d-0,field5d-svd2d-0,field5d-lmafit2d-0}
\end{figure*}

\begin{figure}[htb!]
  \centering
  \subfigure[]{\includegraphics[width=0.9\columnwidth]{field/Fig/z-decimated}
    \label{fig:z-decimated}}
 \subfigure[]{\includegraphics[width=0.9\columnwidth]{field/Fig/z-svd}
    \label{fig:z-svd}}
  \subfigure[]{\includegraphics[width=0.9\columnwidth]{field/Fig/z-lmafit}
    \label{fig:z-lmafit}}
   \caption{Zoomed comparison of reconstruction performance in common offset gathers for the 5D field data example. (a) Observed data with roughly 80\% traces missing. (b) Reconstructed data using the SVD-based low-rank approximation method. (c) Reconstructed data using the SVD-free low-rank approximation method.}
\label{fig:z-decimated,z-svd,z-lmafit}
\end{figure}

@@

\new{\section{Synthetic examples}}
In this section, we use several\new{synthetic}  examples to demonstrate that the proposed SVD-free low-rank approximation method can significantly accelerate the reconstruction of 5D seismic data without loss of accuracy. \new{Although Hankel and Toeplitz matrices \cite{top1,top2} seem to be similar as the low-rank matrix, but in some cases one can lead us to better results in comparison to the other one. According to our experience, the Toeplitz matrix results in a slightly better reconstruction performance. So, here, we prefer to use the Toeplitz matrix as the low-rank matrix.} The first example is a synthetic example with linear events. We first construct a clean synthetic data, then we add some random noise to the data and also randomly remove some traces from the noisy data to simulate the incomplete seismic data. The size of this synthetic data is $100\times10\times10\times10\times10$. A comparison of the clean data, noisy data, and the incomplete data is presented in Figure \ref{fig:syn5d-clean-3d,syn5d-noisy-3d,syn5d-decimated-3d}, from which we can see that the added random noise is so strong that most useful seismic signals are not visible. Since 80\% of the total traces are removed from the original noisy data, the observed data is considered to be highly incomplete. We compare the reconstruction performance between the SVD-based low-rank approximation method and the SVD-free  low-rank approximation method in Figure \ref{fig:syn5d-svd-3d,syn5d-n-svd-3d,syn5d-r-svd-3d,syn5d-lmafit-3d,syn5d-n-lmafit-3d,syn5d-r-lmafit-3d}. The top row in Figure \ref{fig:syn5d-svd-3d,syn5d-n-svd-3d,syn5d-r-svd-3d,syn5d-lmafit-3d,syn5d-n-lmafit-3d,syn5d-r-lmafit-3d} corresponds to the results from the SVD-based low-rank approximation method and the bottom row in Figure \ref{fig:syn5d-svd-3d,syn5d-n-svd-3d,syn5d-r-svd-3d,syn5d-lmafit-3d,syn5d-n-lmafit-3d,syn5d-r-lmafit-3d} corresponds to the results from the SVD-free low-rank approximation method. In  Figure \ref{fig:syn5d-svd-3d,syn5d-n-svd-3d,syn5d-r-svd-3d,syn5d-lmafit-3d,syn5d-n-lmafit-3d,syn5d-r-lmafit-3d}, we show the reconstructed data in common offset gathers in the left column, the removed noise from the noisy data in the middle column, and the reconstruction error in the right column. The results show that the SVD-based and SVD-free low-rank approximation methods both obtain very successful reconstructions. The gaps in the observed incomplete data have been filled with seismic signals for both methods. The removed noise cubes of both methods do not contain obvious signal patterns, indicating that there are no significant damages to the useful signals during the reconstruction. The reconstruction error cubes also demonstrate that the error is negligible for both methods. However, for this example, the proposed SVD-free method is much faster than the SVD-based method. As tested on a Mac Pro Laptop equipped with an Intel Core i7 CPU clocked at 2.5 GHz and 16 GB of RAM, the SVD-based method takes 603 s to finish the calculation while the proposed only takes 153 s. The computational speedup is almost four times. A comparison of the computational efficiency of different methods for different data sizes is shown in Table \ref{tbl:time}.

@@

\begin{thebibliography}{00}
\bibitem{ronen}
J.~Ronen, ``Wave-equation trace interpolation,'' \emph{Geophysics}, vol.~52,
  pp. 973--984, 1987.

\bibitem{spitz1991}
S.~Spitz, ``Seismic trace interpolation in the f-x domain,'' \emph{Geophysics},
  vol.~56, pp. 785--794, 1991.

\bibitem{chen1995}
J.~Chen and J.~Cihlar, ``Quantifying the effect of canopy architecture on
  optical measurements of leaf area index using two gap size analysis
  methods,'' \emph{IEEE Transactions on Geoscience and Remote Sensing},
  vol.~33, pp. 777--787, 1995.

\bibitem{canning1996}
A.~Canning and G.~H.~F. Gardner, ``Reducing 3{D} acquisition footprint for 3{D}
  dmo and 3{D} prestack migration,'' \emph{Geophysics}, vol.~63, pp.
  1177--1183, 1996.

\bibitem{fomel2002pwd}
S.~Fomel, ``Application of plane-wave destruction filters,'' \emph{Geophysics},
  vol.~67, pp. 1946--1960, 2002.

\bibitem{fomel2003}
------, ``Seismic reflection data interpolation with differential offset and
  shot continuation,'' \emph{Geophysics}, vol.~68, pp. 733--744, 2003.

\bibitem{abmapocs}
R.~Abma and N.~Kabir, ``\protect{3D} interpolation of irregular data with a
  \protect{POCS} algorithm,'' \emph{Geophysics}, vol.~71, pp. E91--E97, 2006.

\bibitem{herrmann2008non}
F.~J. Herrmann and G.~Hennenfent, ``Non-parametric seismic data recovery with
  curvelet frames,'' \emph{Geophysical Journal International}, vol. 173, no.~1,
  pp. 233--248, 2008.

\bibitem{protter2009}
M.~Protter and M.~Elad, ``Image sequence denoising via sparse and redundant
  representations,'' \emph{IEEE Trans Image Process}, vol.~18, no.~1, pp.
  27--35, 2009.

\bibitem{mostafa2010}
M.~Naghizadeh and M.~D. Sacchi, ``Beyond alias hierarchical scale curvelet
  interpolation of regularly and irregularly sampled seismic data,''
  \emph{Geophysics}, vol.~75, p. WB189–WB202, 2010.

\bibitem{mostafa2013}
------, ``Multidimensional de-aliased cadzow reconstruction of seismic
  records,'' \emph{Geophysics}, vol.~78, no.~1, pp. A1--A5, 2010.

\bibitem{wang2011recovery}
Y.~Wang, J.~Cao, and C.~Yang, ``Recovery of seismic wavefields based on
  compressive sensing by an l 1-norm constrained trust region method and the
  piecewise random subsampling,'' \emph{Geophysical Journal International},
  vol. 187, no.~1, pp. 199--213, 2011.

\bibitem{yanan2014}
Y.~Tian, Y.~Li, and B.~Yang, ``Variable-eccentricity hyperbolic-trace {TFPF}
  for seismic random noise attenuation,'' \emph{IEEE Transactions on Geoscience
  and Remote Sensing}, vol.~52, pp. 6449--6458, 2014.

\bibitem{yanhui2016}
Y.~Zhou, J.~Gao, W.~Chen, and P.~Frossard, ``Seismic simultaneous source
  separation via patchwise sparse representation,'' \emph{IEEE Transactions on
  Geoscience and Remote Sensing}, vol.~54, pp. 5271--5284, 2016.

\bibitem{kazemi2016}
N.~Kazemi, E.~Bongajum, and M.~D. Sacchi, ``Surface-consistent sparse
  multichannel blind deconvolution of seismic signals,'' \emph{IEEE
  Transactions on Geoscience and Remote Sensing}, vol.~54, pp. 3200--3207,
  2016.
\bibitem{shuwei2016cs}
S.~Gan, S.~Wang, Y.~Chen, X.~Chen, W.~Huang, and H.~Chen, ``Compressive sensing
  for seismic data reconstruction via fast projection onto convex sets based on
  seislet transform,'' \emph{Journal of Applied Geophysics}, vol. 130, pp.
  194--208, 2016.

\bibitem{lorenzi2016}
L.~Lorenzi, F.~Melgani, and G.~Mercier, ``Missing-area reconstruction in
  multispectral images under a compressive sensing perspective,'' \emph{IEEE
  Transactions on Geoscience and Remote Sensing}, vol.~51, pp. 3998--4008,
  2016.

\bibitem{shaohuan2016gap}
S.~Zu, H.~Zhou, Y.~Chen, X.~Pan, S.~Gan, and D.~Zhang, ``Interpolating big gaps
  using inversion with slope constraint,'' \emph{IEEE Geoscience and Remote
  Sensing Letters}, vol.~13, pp. 1369--1373, 2016.

\bibitem{yangkang2016irr5d}
Y.~Chen, D.~Zhang, Z.~Jin, X.~Chen, S.~Zu, W.~Huang, and S.~Gan, ``Simultaneous
  denoising and reconstruction of 5{D} seismic data via damped rank-reduction
  method,'' \emph{Geophysical Journal International}, vol. 206, pp. 1695--1717,
  2016.

\bibitem{amir2017ieee}
R.~Anvari, M.~A.~N. Siahsar, S.~Gholtashi, A.~R. Kahoo, and M.~Mohammadi,
  ``Seismic random noise attenuation using synchrosqueezed wavelet transform
  and low-rank signal matrix approximation,'' \emph{IEEE Transactions on
  Geoscience and Remote Sensing}, vol.~55, no.~11, pp. 6574--6581, 2017.

\bibitem{verschuure1991}
D.~J. Verschuur, ``Surface-related multiple elimination: An inversion
  approach,'' \emph{PhD thesis, Delft University of Technology}, 1991.

>>>>

%\history{Date of publication xxxx 00, 0000, date of current version xxxx 00, 0000.}
%\doi{10.1109/ACCESS.2020.DOI}

\DeclareRobustCommand{\dlo}[1]{}
\DeclareRobustCommand{\wen}[1]{#1}

\published{IEEE Access, 8, 175501-175512, (2020)}

@@

%\corresp{Corresponding author: M. Bai (e-mail: baimin2016@126.com).}
%\tfootnote{This research is sponsored by the National Natural Science Foundation of China (Nos. \new{41904110,}41704121), \new{the China Postdoctoral Science Foundation (No. 2019M662567), and the Postdoctoral Science Foundation of Hubei Province.}}
%

%\def\thevol{8}
%\def\myyear{2020}

@@

%\begin{keywords}
%Multidimensional seismic data, \old{lowrank}\new{low-rank} approximation, seismic data processing, \new{seismic reconstruction, matrix completion}
%\end{keywords}

%\titlepgskip=-15pt

@@

\inputdir{linear}
\multiplot{3}{syn5d-clean-3d,syn5d-noisy-3d,syn5d-decimated-3d}{width=0.3\columnwidth}{Common offset gathers for the 5D synthetic example with linear events. (a) Clean data. (b) Noisy data. (c) The observed data with 80\% traces randomly removed.}

\multiplot{6}{syn5d-svd-3d,syn5d-n-svd-3d,syn5d-r-svd-3d,syn5d-lmafit-3d,syn5d-n-lmafit-3d,syn5d-r-lmafit-3d}{width=0.3\columnwidth}{Comparison of reconstruction performance in common offset gathers for the 5D synthetic example with linear events. (a) Reconstructed data using the SVD-based low-rank approximation method. (b) The removed noise from the noisy data corresponding to (a). (c) The reconstruction error corresponding to (a). (d) Reconstructed data using the SVD-free low-rank approximation method. (e) The removed noise from the noisy data corresponding to (d). (f) The reconstruction error corresponding to (d).}

\multiplot{3}{syn5d-clean-3d-1,syn5d-noisy-3d-1,syn5d-decimated-3d-1}{width=0.3\columnwidth}{Common midpoint gathers for the 5D synthetic example with linear events. (a) Clean data. (b) Noisy data. (c) The observed data with 80\% traces randomly removed.}

\multiplot{6}{syn5d-svd-3d-1,syn5d-n-svd-3d-1,syn5d-r-svd-3d-1,syn5d-lmafit-3d-1,syn5d-n-lmafit-3d-1,syn5d-r-lmafit-3d-1}{width=0.3\columnwidth}{Comparison of reconstruction performance in common midpoint gathers for the 5D synthetic example with linear events. (a) Reconstructed data using the SVD-based low-rank approximation method. (b) The removed noise from the noisy data corresponding to (a). (c) The reconstruction error corresponding to (a). (d) Reconstructed data using the SVD-free low-rank approximation method. (e) The removed noise from the noisy data corresponding to (d). (f) The reconstruction error corresponding to (d).}

\multiplot{5}{syn5d-clean-2d,syn5d-noisy-2d,syn5d-decimated-2d,syn5d-svd-2d,syn5d-lmafit-2d}{width=0.3\columnwidth}{Comparison of reconstruction performance in a 2D section. (a) Clean data. (b) Noisy data. (c) The observed data. (d) Reconstructed data using the SVD-based low-rank approximation method. (e) Reconstructed data using the SVD-free low-rank approximation method.}

\plot{synth-ss}{width=0.3\columnwidth}{Comparison of the middle trace amplitude of each cubes in different datasets. The black line denotes the trace from the clean data.  The red line denotes the noisy data. The observed trace cannot be seen from this FIGURE because it is a blank trace in the selection position. The SVD-based and SVD-free methods are denoted as blue and green lines, respectively.}

\multiplot{4}{simi-syn5d-noisy-2d,simi-syn5d-decimated-2d,simi-syn5d-svd-2d,simi-syn5d-lmafit-2d}{width=0.45\columnwidth}{Comparison of reconstruction performance in terms of local similarity. (a) Local similarity of the noisy data. (b) Local similarity of the observed data. (c) Local similarity of the reconstructed data using SVD-based low-rank approximation method. (d) Local similarity of the reconstructed data using SVD-free low-rank approximation method.}

\inputdir{curve}
\multiplot{3}{curve5d-clean-3d-1,curve5d-noisy-3d-1,curve5d-decimated-3d-1}{width=0.3\columnwidth}{Common midpoint gathers for the 5D synthetic example with hyperbolic events. (a) Clean data. (b) Noisy data. (c) The observed data with 80\% traces randomly removed.}

\multiplot{6}{curve5d-svd-3d-1,curve5d-n-svd-3d-1,curve5d-r-svd-3d-1,curve5d-lmafit-3d-1,curve5d-n-lmafit-3d-1,curve5d-r-lmafit-3d-1}{width=0.3\columnwidth}{Comparison of reconstruction performance in common midpoint gathers for the 5D synthetic example with hyperbolic events. (a) Reconstructed data using the SVD-based low-rank approximation method. (b) The removed noise from the noisy data corresponding to (a). (c) The reconstruction error corresponding to (a). (d) Reconstructed data using the SVD-free low-rank approximation method. (e) The removed noise from the noisy data corresponding to (d). (f) The reconstruction error corresponding to (d).}

\multiplot{4}{simi-curve5d-noisy-3d-1,simi-curve5d-decimated-3d-1,simi-curve5d-svd-3d-1,simi-curve5d-lmafit-3d-1}{width=0.45\columnwidth}{Comparison of reconstruction performance in terms of local similarity. (a) Local similarity of the noisy data. (b) Local similarity of the observed data. (c) Local similarity of the reconstructed data using SVD-based low-rank approximation method. (d) Local similarity of the reconstructed data using SVD-free low-rank approximation method.}

\inputdir{./}
\multiplot{3}{field5d-decimated2d-0,field5d-svd2d-0,field5d-lmafit2d-0}{width=0.8\columnwidth}{Comparison of reconstruction performance in common offset gathers for the 5D field data example. (a) Observed data with roughly 80\% traces missing. (b) Reconstructed data using the SVD-based low-rank approximation method. (c) Reconstructed data using the SVD-free low-rank approximation method. The frame box areas are zoomed and highlighted in FIGURE \ref{fig:z-decimated,z-svd,z-lmafit}.}

\multiplot{3}{z-decimated,z-svd,z-lmafit}{width=0.45\columnwidth}{Zoomed comparison of reconstruction performance in common offset gathers for the 5D field data example. (a) Observed data with roughly 80\% traces missing. (b) Reconstructed data using the SVD-based low-rank approximation method. (c) Reconstructed data using the SVD-free low-rank approximation method.}

@@

\new{\section{Synthetic examples}}
In this section, we use several\new{synthetic}  examples to demonstrate that the proposed SVD-free low-rank approximation method can significantly accelerate the reconstruction of 5D seismic data without loss of accuracy. \new{Although Hankel and Toeplitz matrices seem to be similar as the low-rank matrix, but in some cases one can lead us to better results in comparison to the other one. According to our experience, the Toeplitz matrix results in a slightly better reconstruction performance. So, here, we prefer to use the Toeplitz matrix as the low-rank matrix.} The first example is a synthetic example with linear events. We first construct a clean synthetic data, then we add some random noise to the data and also randomly remove some traces from the noisy data to simulate the incomplete seismic data. The size of this synthetic data is $100\times10\times10\times10\times10$. A comparison of the clean data, noisy data, and the incomplete data is presented in Figure \ref{fig:syn5d-clean-3d,syn5d-noisy-3d,syn5d-decimated-3d}, from which we can see that the added random noise is so strong that most useful seismic signals are not visible. Since 80\% of the total traces are removed from the original noisy data, the observed data is considered to be highly incomplete. We compare the reconstruction performance between the SVD-based low-rank approximation method and the SVD-free  low-rank approximation method in Figure \ref{fig:syn5d-svd-3d,syn5d-n-svd-3d,syn5d-r-svd-3d,syn5d-lmafit-3d,syn5d-n-lmafit-3d,syn5d-r-lmafit-3d}. The top row in Figure \ref{fig:syn5d-svd-3d,syn5d-n-svd-3d,syn5d-r-svd-3d,syn5d-lmafit-3d,syn5d-n-lmafit-3d,syn5d-r-lmafit-3d} corresponds to the results from the SVD-based low-rank approximation method and the bottom row in Figure \ref{fig:syn5d-svd-3d,syn5d-n-svd-3d,syn5d-r-svd-3d,syn5d-lmafit-3d,syn5d-n-lmafit-3d,syn5d-r-lmafit-3d} corresponds to the results from the SVD-free low-rank approximation method. In  Figure \ref{fig:syn5d-svd-3d,syn5d-n-svd-3d,syn5d-r-svd-3d,syn5d-lmafit-3d,syn5d-n-lmafit-3d,syn5d-r-lmafit-3d}, we show the reconstructed data in common offset gathers in the left column, the removed noise from the noisy data in the middle column, and the reconstruction error in the right column. The results show that the SVD-based and SVD-free low-rank approximation methods both obtain very successful reconstructions. The gaps in the observed incomplete data have been filled with seismic signals for both methods. The removed noise cubes of both methods do not contain obvious signal patterns, indicating that there are no significant damages to the useful signals during the reconstruction. The reconstruction error cubes also demonstrate that the error is negligible for both methods. However, for this example, the proposed SVD-free method is much faster than the SVD-based method. As tested on a Mac Pro Laptop equipped with an Intel Core i7 CPU clocked at 2.5 GHz and 16 GB of RAM, the SVD-based method takes 603 s to finish the calculation while the proposed only takes 153 s. The computational speedup is almost four times. A comparison of the computational efficiency of different methods for different data sizes is shown in Table \ref{tbl:time}.

@@

\bibliographystyle{seg}
\bibliography{5d}

%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a1.png}}]{First A. Author} (M'76--SM'81--F'87) and all authors may include 
%biographies. Biographies are often not included in conference-related
%papers. This author became a Member (M) of IEEE in 1976, a Senior
%Member (SM) in 1981, and a Fellow (F) in 1987. The first paragraph may
%contain a place and/or date of birth (list place, then date). Next,
%the author's educational background is listed. The degrees should be
%listed with type of degree in what field, which institution, city,
%state, and country, and year the degree was earned. The author's major
%field of study should be lower-cased. 
%
%The second paragraph uses the pronoun of the person (he or she) and not the 
%author's last name. It lists military and work experience, including summer 
%and fellowship jobs. Job titles are capitalized. The current job must have a 
%location; previous positions may be listed 
%without one. Information concerning previous publications may be included. 
%Try not to list more than three books or published articles. The format for 
%listing publishers of a book within the biography is: title of book 
%(publisher name, year) similar to a reference. Current and previous research 
%interests end the paragraph. The third paragraph begins with the author's 
%title and last name (e.g., Dr.\ Smith, Prof.\ Jones, Mr.\ Kajor, Ms.\ Hunter). 
%List any memberships in professional societies other than the IEEE. Finally, 
%list any awards and work for IEEE committees and publications. If a 
%photograph is provided, it should be of good quality, and 
%professional-looking. Following are two examples of an author's biography.
%\end{IEEEbiography}
%
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a2.png}}]{Second B. Author} was born in Greenwich Village, New York, NY, USA in 
%1977. He received the B.S. and M.S. degrees in aerospace engineering from 
%the University of Virginia, Charlottesville, in 2001 and the Ph.D. degree in 
%mechanical engineering from Drexel University, Philadelphia, PA, in 2008.
%
%From 2001 to 2004, he was a Research Assistant with the Princeton Plasma 
%Physics Laboratory. Since 2009, he has been an Assistant Professor with the 
%Mechanical Engineering Department, Texas A{\&}M University, College Station. 
%He is the author of three books, more than 150 articles, and more than 70 
%inventions. His research interests include high-pressure and high-density 
%nonthermal plasma discharge processes and applications, microscale plasma 
%discharges, discharges in liquids, spectroscopic diagnostics, plasma 
%propulsion, and innovation plasma applications. He is an Associate Editor of 
%the journal \emph{Earth, Moon, Planets}, and holds two patents. 
%
%Dr. Author was a recipient of the International Association of Geomagnetism 
%and Aeronomy Young Scientist Award for Excellence in 2008, and the IEEE 
%Electromagnetic Compatibility Society Best Symposium Paper Award in 2011. 
%\end{IEEEbiography}
%
